---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "40%"
)

library(relper)

```
# relper

The goal of relper is to provide miscellaneous functions to assist in data cleaning and visualization.

## Installation

You can install the developer version of relper with:

``` r
remotes::install_github("vbfelix/relper")
```

## ggplot2 functions

```{r example}
library(ggplot2)

plot <- 
ggplot(mtcars,aes(drat,hp))+
  geom_point()

plot
```

### theme_x

```{r}
plot + theme_x()
```

### theme_y

```{r}
plot + theme_y()
```

### theme_map

```{r}
plot + theme_map()
```

### flip_y_title

```{r}
plot + flip_y_title
```

### water_mark

```{r}
url <- "https://upload.wikimedia.org/wikipedia/commons/thumb/1/1b/R_logo.svg/1200px-R_logo.svg.png"

logo <- water_mark(url, local_file = F)

plot + annotation_custom(logo)

plot + annotation_custom(logo, xmin = 4.75, xmax = 5, ymin = 315, ymax = 335)
```

## Metrics

```{r}
x <- rexp(20,.5)

x
```


### Coefficient of Variation (CV)

```{r}
#raw
cv(x, perc = F)

#%
cv(x, perc = T)
```

### Mean's

```{r}
num_mean(x)
```

#### Harmonic mean

```{r}
harmonic_mean(x)
```

#### Geometric mean

```{r}
geometric_mean(x)
```

### Numeric univariate summary statistics

```{r}
num_summary(x)
```

### Correlations

```{r}
y <- rnorm(20)

num_corr(x,y)
```

## Others

### Area under the curve

```{r aux example}
x <- seq(-3,3,l = 100)

y <- dnorm(x)

plot(x,y)
```

```{r}
#from min to max of x
range(x)

auc(x,y)

#from -2 to 2
auc(x,y,limits = c(-2,2))

#from -1 to 1
auc(x,y,limits = c(-1,1))

```





