---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(relper)

```
# relper

The goal of relper is to provide miscellaneous functions to assist in data cleaning and preparation

## Installation

You can install the developer version of relper with:

``` r
remotes::install_github("vbfelix/relper")
```

## ggplot2 functions

This is a basic example which shows you how to solve a common problem:

```{r example}
library(ggplot2)
plot <- 
ggplot(mtcars,aes(drat,hp))+
  geom_point()

plot
```

### theme_x

```{r}
plot + theme_x()
```

### theme_y

```{r}
plot + theme_y()
```

### theme_x

```{r}
plot + theme_map()
```


## metrics

```{r}
set.seed(123)
x <- rexp(20,.5)

x

```


### Coefficient of Variation (CV)

```{r}

#raw
cv(x, perc = F)

#%
cv(x, perc = T)

```

### Harmonic mean

```{r}

harmonic_mean(x)

```

## Others

### Area under the curve

```{r aux example}
x <- seq(-3,3,l = 100)
y <- dnorm(x)
plot(x,y)
```

```{r}
auc(x,y)
```





