---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  cache = T,
  message = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "40%"
)

set.seed(333)
  
library(relper)

```

# relper

The goal of relper is to provide miscellaneous functions to assist in data cleaning and visualization.

## Installation

You can install the developer version of relper with:

``` r
remotes::install_github("vbfelix/relper")
```

## ggplot2 functions

```{r example}
library(ggplot2)

plot <- 
ggplot(mtcars,aes(drat,hp))+
  geom_point()

plot
```

### theme_x

```{r}
plot + theme_x()
```

### theme_y

```{r}
plot + theme_y()
```

### theme_map

```{r}
plot + theme_map()
```

### flip_y_title

```{r}
plot + flip_y_title
```

### water_mark

```{r}
url <- "https://upload.wikimedia.org/wikipedia/commons/thumb/1/1b/R_logo.svg/1200px-R_logo.svg.png"

logo <- water_mark(url, local_file = F)

plot + annotation_custom(logo)

plot + annotation_custom(logo, xmin = 4.65, xmax = Inf, ymin = 310, ymax = Inf)
```

### calendar data and plot (from ggcal)

```{r}
library(tidyverse)

date <- seq(as.Date("2016-01-01"), as.Date("2016-02-29"),by = "1 day")

#data
df <-
  tibble(dates = rep(date, 3),
         value = num_format(rnorm(3*length(date)),0),
         grp = rep(c("A","B","C"), each = length(date))) %>%
  nest(-grp) %>%
  mutate(aux = map(data,~cal_data(.$dates,.$value))) %>%
  select(-data) %>%
  unnest() 

head(df)

df %>% 
  cal_plot(facet = grp)+
  scale_fill_viridis_d()
```

### Duas axis plot

```{r}
x <- seq(0,1,l = 100)

y <- cumsum(rnorm(100))

z <- y^3 + rnorm(100)

df <- data.frame(x,y,z)

dual_plot(df,x_axis = "x",y_left = "y",y_right = "z")
```


## Metrics

```{r}
x <- rexp(20,.5)

x
```


### Coefficient of Variation (CV)

```{r}
#raw
cv(x, perc = F)

#%
cv(x, perc = T)
```

### Mean's

```{r}
num_mean(x)
```

#### Harmonic mean

```{r}
harmonic_mean(x)
```

#### Geometric mean

```{r}
geometric_mean(x)
```

### Numeric univariate summary statistics

```{r}
num_summary(x)
```

### Correlations

```{r}
y <- rnorm(20)

num_corr(x,y)
```

## Others

### Scaling

```{r}
x <- seq(-3,3,l = 10)

x

y <- scale01(x)

y

plot(x,y)

z <- scale01(x,lim_sup = 100)

plot(x,z)
```

### Area under the curve

```{r aux example}
x <- seq(-3,3,l = 100)

y <- dnorm(x)

plot(x,y)
```

```{r}
#from min to max of x
range(x)

auc(x,y)

#from -2 to 2
auc(x,y,limits = c(-2,2))

#from -1 to 1
auc(x,y,limits = c(-1,1))

```





