---
title: "calc_ functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{calc_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE}
library(relper)
library(dplyr)
library(ggplot2)
```

`calc_` functions compute a certain value.

# calc_acf

The goal of `calc_acf` is to compute the auto-correlation function, given by:

$$\frac{\sum_\limits{t = k+1}^{n}(x_t - \bar{x})(x_{t-k} - \bar{x})}{\sum_\limits{t = 1}^{n} (x_t - \bar{x})^2 },$$
where:

 - $x_t$ is a time series of length $n$;
 - $x_{t-k}$ is a shifted time series by $k$ units in time;
 - $\bar{x}$ is the average of the time series.

```{r calc_acf}
x <- rnorm(100)

calc_acf(x)

```

If you pass a second vector in the argument `y` the cross-correlation will be computed instead:

$$\frac{n \left( \sum_\limits{t = 1}^{n}x_ty_t \right) - \left[\left(\sum_\limits{t = 1}^{n}x_t \right) \left(\sum_\limits{t = 1}^{n}y_t\right) \right]}{\sqrt{\left[n \left( \sum_\limits{t = 1}^{n}x_t^2 \right) - \left( \sum_\limits{t = 1}^{n}x_t \right)^2\right]\left[n \left( \sum_\limits{t = 1}^{n}y_t^2 \right) - \left( \sum_\limits{t = 1}^{n}y_t \right)^2\right]}},$$
where:

 - $x_t$ is a time series of length $n$;
 - $y_t$ is a time series of length $n$.

```{r calc_acf-ccf}
y <- rexp(100)

calc_acf(x,y)

```

# calc_auc

The goal of `calc_auc` is to compute the area under a curve (AUC).

```{r calc_auc-data}
x <- seq(-3,3,l = 100)

y <- dnorm(x)
```

```{r calc_auc-base-plot, echo = F}
ggplot(tibble(x = x,
              y = y),
       aes(x,y))+
  geom_point()+
  plt_theme_y()
```

The function default compute the area considering the range of `x`.

```{r calc_auc-99}
#from min to max of x
range(x)

calc_auc(x,y)
```

```{r calc_auc-plot-99, echo = F}
df_auc <-
  tibble(
    x = x,
    y = y)

df_auc %>% 
ggplot(aes(x,y))+
  geom_area(alpha = .7, fill = "chocolate2")+
  geom_point()+
  plt_theme_y()+
  annotate("text",
           x = mean(x),y = mean(y),label = round(calc_auc(x,y),3),
           fontface = "bold", size = 7)+
  geom_vline(xintercept = c(-3,3), linetype = "dashed")+
  scale_x_continuous(breaks = -5:5)+
  scale_y_continuous(expand = c(0,0))
```

But you can define the argument `limits` to get the AUC of that respective range.

```{r calc_auc-95}
#from -2 to 2
calc_auc(x,y,limits = c(-2,2))
```

```{r calc_auc-plot-95, echo = F}
df_auc %>% 
  ggplot(aes(x,y))+
  geom_area(data = df_auc %>% 
              filter(between(x,-2,2)),
              alpha = .7, fill = "chocolate2")+
  geom_point()+
  plt_theme_y()+
  annotate("text",
           x = mean(x),y = mean(y),label = round(calc_auc(x,y,limits = c(-2,2)),3),
           fontface = "bold", size = 7)+
  geom_vline(xintercept = c(-2,2), linetype = "dashed")+
  scale_x_continuous(breaks = -5:5)+
  scale_y_continuous(expand = c(0,0))
```

```{r calc_auc-68}
#from -1 to 1
calc_auc(x,y,limits = c(-1,1))
```

```{r calc_auc-plot-68, echo = F}
df_auc %>% 
  ggplot(aes(x,y))+
  geom_area(data = df_auc %>% 
              filter(between(x,-1,1)),
              alpha = .7, fill = "chocolate2")+
  geom_point()+
  plt_theme_y()+
  annotate("text",
           x = mean(x),y = mean(y),label = round(calc_auc(x,y,limits = c(-1,1)),3),
           fontface = "bold", size = 7)+
  geom_vline(xintercept = c(-1,1), linetype = "dashed")+
  scale_x_continuous(breaks = -5:5)+
  scale_y_continuous(expand = c(0,0))
```

# calc_corr
The goal of `calc_corr` is to compute Pearson, Kendall and Spearman correlation coefficients.

```{r calc_corr}
x <- rnorm(100)

y <- rnorm(100)

calc_corr(x,y)
```


# calc_cramers_v

The goal of `calc_cramers_v` is to compute Cramer's V, given by:

$$\sqrt{\frac{X^2}{n\min(r-1,c-1)}},$$
where:

  - $X^2$ the chi-square statistic;
  - $n$ is the sample size;
  - $r$ is the number of rows in the contingency table;
  - $c$ is the number of columns in the contingency table.

```{r calc_cramers_v}
m <- matrix(c(12, 5, 7, 7), ncol = 2)

chi_square <- chisq.test(m)

calc_cramers_v(chi_square)
```

# calc_cv
The goal of `calc_cv` is to compute the coefficient of variation (CV), given by:

$$\frac{s}{\bar{x}},$$
where:

 - $s$ is the sample standard deviation;
 - $\bar{x}$ is the sample mean.
 
```{r calc_cv}
set.seed(123);x <- rexp(n = 100)

calc_cv(x)
```

If you set the argument `as_perc` to `TRUE`, the CV will be multiplied by 100.

```{r calc_cv-as-perc}
calc_cv(x,as_perc = TRUE)
```

# calc_date_aux

The goal of `calc_date_aux` is to compute variables derived from date, such as year, month, day, etc.

```{r calc_date_aux}
dt <- seq(as.Date("1910/1/1"), as.Date("1911/1/1"), "days")

df_dt <- data.frame(dt = dt)

calc_date_aux(df_dt,dt) %>% glimpse()
```

The new variables respective to the date are:

 - **mon_abb:** a factor with the abbreviated name of the month;
 - **mon_lbl:** a factor with the complete name of the month;
 - **mon_num:** the number of the month;
 - **day_num:** the number of the day within the month;
 - **year_num:** the number of the year;
 - **year_lbl:** a factor with the number of the year;
 - **week_num:** the number of complete seven day periods that have occurred between the date and January 1st, plus one;
 - **epiweek_num:** the US CDC version of epidemiological week. Starts on Sunday;
 - **isoweek_num:**  the week as it would appear in the ISO 8601 system, which uses a reoccurring leap week. Starts on Monday;
 - **wday_abb:** a factor with the week day abbreviated name;
 - **wday_lbl:** a factor with the week day complete name.
 
# calc_date_diff

The goal of `calc_date_diff` is to compute the difference between two dates.

```{r calc_date_diff}
date1 <- lubridate::dmy("01/05/1998")

date2 <- lubridate::dmy("21/11/2018")

calc_date_diff(date1 = date1,date2 = date2,unit = "days")
```

So, the difference in days of 21/11/2018 and 01/05/1998 is 7,509 days.

If you need to add a constant to your difference you can use the argument `add`.

```{r calc_date_diff-add}
calc_date_diff(date1 = date1,date2 = date2,unit = "days",add = 1)
```

# calc_date_range

The goal of `calc_date_range` is to compute the range of a date vector, and create a string with the minimal and maximum date.

```{r calc_date_range}
dt <- seq(as.Date("1910/1/1"), as.Date("1911/1/1"), "days")

calc_date_range(dt)
```


# calc_geometric_mean

The goal of `calc_geometric_mean` is to compute the geometric mean, given by:

$$\sqrt[n]{\prod\limits_{i=1}^{n}x_i} = \sqrt[n]{x_1\times x_2 \times...\times x_n},$$
where:

 - $x_i$ is a numeric vector of length $n$.

```{r calc_geometric_mean}
calc_geometric_mean(x)
```

# calc_harmonic_mean

The goal of `calc_harmonic_mean` is to compute the harmonic mean, given by:

$$\frac{n}{\sum\limits_{i=1}^{n}\frac{1}{x_i}},$$
where:

 - $x_i$ is a numeric vector of length $n$.

```{r calc_harmonic_mean}
calc_harmonic_mean(x)
```

# calc_mean

The goal of `calc_mean` is to compute the arithmetic, geometric and harmonic mean.

```{r calc_mean}
calc_mean(x)
```

# calc_mode

The goal of `calc_mode` is to compute the mode.

```{r calc_mode-data}
set.seed(123);cat_var <- sample(letters,100,replace = TRUE)

table(cat_var)

```
We can see that the letter "*y*" is the one that appears the most, so it is the mode of the variable.

```{r calc_mode}
calc_mode(cat_var)
```

# calc_peak_density
The goal of `calc_peak_density` is to compute the peak density value of a numeric value.

```{r calc_peak_density-base-plot,echo = F}

pd_plot <-
  ggplot2::ggplot(data = dplyr::tibble(x = x),
                  ggplot2::aes(x))+
  ggplot2::geom_density()+
  relper::plt_theme_y()+
  ggplot2::scale_x_continuous(breaks = 0:20, expand = c(0,0))+
  plt_flip_y_title+
  labs(y = "Density (x)")

pd_plot

```

Let's say we eant to discover what value is the peak in the density above.

```{r calc_peak_density}
calc_peak_density(x)
```

```{r calc_peak_density-plot,echo = F, warning = FALSE, message = FALSE}
x_peak <- relper::calc_peak_density(x)

pd_plot+
  ggplot2::geom_vline(
    xintercept = relper::calc_peak_density(x),
    col = "royalblue4",
    size = 1
    )+
  ggplot2::scale_x_continuous(
    breaks = 0:20,
    expand = c(0,0),
    sec.axis = sec_axis(~.,
                        breaks = x_peak,
                        labels = format_num(x_peak,digits = 3)
                        )
    )

```
