---
title: "format_functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{format_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(relper)
```

`format_` functions modify an existing variable for aesthetic reasons.

# format_digit

The goal of `format_digit` is to add zero(s) on the left of a number.

```{r format_digit}

x <- c(1,4,10,12,100,2000)

format_digit(x)

```

You can also set the argument `digits` to add more zeros.

```{r format_digit-digits}

format_digit(x,digits = 4)

```

# format_num

The goal of `format_num` is to add markers to a number.

```{r format_num}
format_num(12345.67)
```

You can also change the markers for other characters of your choice.

```{r format_num-markers}
format_num(12345.67,decimal_mark = "*",thousand_mark = "#")
```
Also, if you are Brazilian like me, I made a argument to make our lives easier.

```{r format_num-br}
format_num(12345.67,br_mark = TRUE)
```


# format_scale
The goal of `format_scale` is to reescale a variable, given by:

$$y_i = (y_{\max} - y_{\min}) * \left[\frac{x_i - \min(x_i)}{\max(x)-\min(x)} \right] + y_{\min},$$
where:

  - $y_i$ is the reescaled numeric vector;
  - $y_{\max}$ is the new maximum value, after the reescale;
  - $y_{\min}$ is the new minimum value, after the reescale;
  - $x_i$ is the original numeric vector;
  - $\min(x_i)$ is the original minimum value of $x_i$;
  - $\max(x_i)$ is the original maximum value of $x_i$.
  
By default the vector will be reescaled to have 0 and 1 as their new minimum and maximum values.  

```{r format_scale}

x <- seq(-3,3,l = 10)

x

y <- format_scale(x)

y
```

```{r format_scale-plot, echo = F}
plot(x,y)
```

You can also change the range of the new scale.

```{r format_scale-range}

z <- format_scale(x,new_min = 25,new_max = 100)

```

```{r format_scale-range-plot, echo = F}

plot(x,z)

```

# format_p_value

The goal of `format_p_value` is to change a p value, by considering a lower bound value where if is the valor is smaller than it, let's say 0.001, the p value will be changed to *<0.001*

```{r format_p_value}
format_p_value(c(.001,.00000001),lower_bound = 0.001)
```
