---
title: "Functions plt_"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions plt_}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE}
library(relper)
library(ggplot2)

```

`plt_` functions are extensions to `ggplot2` objects. So let's define a base plot to apply the functions.

```{r}
plot <- 
ggplot(mtcars,aes(qsec,mpg))+
  geom_point()

plot
```

# plt_flip_y_title

The goal of `plt_flip_y_title` is to flip the title from y axis, so it can be read horizontally.

```{r plt_flip_y_title}
plot + plt_flip_y_title
```

# plt_identity_line

The goal of `plt_identity_line` is to add a identity line to a scatter plot.

```{r plt_identity_line}
plot + plt_identity_line(color = "red")
```

# plt_no_background

The goal of `plt_no_background` is to remove the background.

```{r plt_no_background}
plot + plt_no_background
```

# plt_no_labels

The goal of `plt_no_labels` is to remove all labels.

```{r plt_no_labels}
plot + plt_no_labels
```

# plt_regression_line

The goal of `plt_regression_line` is to add a regression line to a scatter plot.

```{r plt_regression_line}
plot + plt_regression_line(color = "red")
```

# plt_scale_auto

The goal of `plt_scale_auto` is to add a automatic scale.

```{r plt_scale_auto}
plot + plt_scale_auto(axis = "x",n_breaks = 5)
```

```{r plt_scale_auto-x-y}
plot +
  plt_scale_auto(axis = "xy",n_breaks = 5)
```

# plt_scale_y_mirror

The goal of `plt_scale_y_mirror` is to mirror the y axis on the right side.

```{r plt_scale_y_mirror}
plot_breaks <- seq(0,35,3)

plot_labels <- format_num(plot_breaks,digits = 2)

plot + plt_scale_y_mirror(breaks = plot_breaks, labels = plot_labels)
```

# plt_theme_map

The goal of `plt_theme_map` is to add a theme appropriate for a map.

```{r plt_theme_map}
plot + plt_theme_map()
```

# plt_theme_x

The goal of `plt_theme_x` is to remove major and minor grid lines from y axis.

```{r plt_theme_x}
plot + plt_theme_x(margin = 1)
```

# plt_theme_xy

The goal of `plt_theme_xy` is to remove minor grid lines from x and y axis.

```{r plt_theme_xy}
plot + plt_theme_xy(margin = 2)
```

# plt_theme_y

The goal of `plt_theme_y` is to remove major and minor grid lines from x axis.

```{r plt_theme_y}
plot + plt_theme_y()
```

# plt_water_mark

The goal of `plt_water_mark` is to add a image as a watermark in a `ggplot2` object.

```{r plt_water_mark}
plot + plt_water_mark(vfx_watermark)
```

