---
title: "Functions str_"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions str_}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE}
library(relper)
```

`str_` functions manipulate strings.

# str_clean

The goal of `str_clean` is to remove punctuation and/or accent from a string.

```{r str_clean}
string <- "a..;éâ...íõ"
```

By default the function  will remove punctuation and accent symbols.
```{r default}
str_clean(string)
```

But you can remove only remove the accent symbols.

```{r remove_accent}
str_clean(string,remove_accent = TRUE,remove_punct = FALSE)
```

Also you can remove only remove the accent symbols.

```{r remove_punct}
str_clean(string,remove_accent = FALSE,remove_punct = TRUE)
```

And there is the possibility to substitute the punctuation symbols.

```{r str_clean_punct}
str_clean(string,remove_accent = FALSE,remove_punct = TRUE,sub_punct = "-")
```

# str_extract_char

The goal of `str_extract_char` is to extract a specific character from a string.

```{r str_extract_char}
str_extract_char(string = "abcdef",char = 2)
```
# str_keep

The goal of `str_keep` is to keep only a type of character in the string.

```{r str_keep}
string <- "1Aa45Z89$$%#"
```

By default the function will keep only letters.

```{r text}
str_keep(string,keep = "text")
```

But you can also keep only numbers.

```{r numbers}
str_keep(string,keep = "numbers")
```

And there is also an option  to keep special characters.

```{r special}
str_keep(string,keep = "special")
```

# str_select

The goal of `str_select` is to select part of a string, before, after or between patterns.

```{r str_select}
string <- "begin STRING1 TARGET STRING2 end"
```

By setting the argument `before` you select only the part of the string before this pattern.

```{r before}
str_select(string,before = "STRING2")
```

By setting the argument `after` you select only the part of the string after this pattern.

```{r after}
str_select(string,after = "STRING1")
```

By setting the arguments `brfore` and `after` you select only the part of the string between both patterns.

```{r both}
str_select(string,"STRING1","STRING2")
```

# str_to_abb

The goal of `str_to_abb` is to apply uppercase to strings with a number of characters lower than the argument `n_abb` (default = 3), and only uppercase the first letter for the others words, so abbreviations will not be affected. 

```{r str_to_abb}
abcde_string <- c("aaaaa","bb","ccc","dddd")

str_to_abb(abcde_string,n_abb = 3)
```
