---
title: "Functions palette_"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions palette_}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  # collapse = TRUE,
  fig.width = 10,warning = FALSE,message = FALSE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE, echo = FALSE}
library(relper)

palette_vignette <-
  function(func_palette){
    
    palettes <- as.character(unlist(formals(func_palette)$name)) 
    palettes <- palettes[-1]
    names(palettes) <- palettes
    
    palettes_df <-
      tibble::tibble(palette = palettes) %>% 
      dplyr::mutate(color = purrr::map(palettes, func_palette)) %>% 
      tidyr::unnest(cols = c(color)) %>% 
      dplyr::group_by(palette) %>% 
      dplyr::mutate(id = dplyr::row_number()) %>% 
      dplyr::mutate(x = forcats::fct_reorder(.f = color,.x = id))
    
    colors <- palettes_df$color
    names(colors) <- colors
    
    palettes_df %>% 
      ggplot2::ggplot(ggplot2::aes(x,palette, fill = color))+
      ggplot2::geom_tile(col = "white",size = 1)+
      ggplot2::scale_fill_manual(values = colors)+
      ggplot2::scale_x_discrete(expand = c(0,0))+
      ggplot2::scale_y_discrete(expand = c(0,0))+
      ggplot2::theme(
        legend.position = "none",
        axis.ticks = ggplot2::element_blank(),
        axis.text.y = ggplot2::element_blank(),
        strip.background = ggplot2::element_blank(),
        strip.text = ggplot2::element_text(face = "bold",size = 15),
        axis.text.x = ggplot2::element_text(color = "black", size = 10)
      )+
      ggplot2::facet_wrap(~palette,scales = "free", ncol = 1)+
      relper::plt_no_labels
  }
```

`palette_` functions provides color schemes that are influenced by popular culture media, such as: TV series, movies, video games, animations, etc.

# palette_div

The goal of `palette_div` is to return a palette for a divergent scale, with the following palettes:


```{r, echo = FALSE,fig.height = 14}

palette_vignette(func_palette = palette_div)

```

# palette_qua

The goal of `palette_qua` is to return a palette for a qualitative scale, with the following palettes:

```{r, echo = FALSE,fig.height = 18}

palette_vignette(func_palette = palette_qua)

```

# palette_seq

The goal of `palette_seq` is to return a palette for a sequential scale, with the following palettes:

```{r, echo = FALSE,fig.height = 17}

palette_vignette(func_palette = palette_seq)

```


