---
title: "Functions dttm_"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions dttm_}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE}
library(relper)
```

`dttm_` functions help in the manipulation of date and datetime variables.

# dttm_diff

The goal of `dttm_diff` is to compute the difference between two dates.

```{r dttm_diff}
date1 <- lubridate::dmy("01/05/1998")

date2 <- lubridate::dmy("21/11/2018")

dttm_diff(date1 = date1,date2 = date2,unit = "days")
```

So the time difference between 21/11/2018 and 01/05/1998 is 7,509 days.

If you need to add a constant to your difference, use the `add` argument.

```{r dttm_diff-add}
dttm_diff(date1 = date1,date2 = date2,unit = "days",add = 1)
```


# dttm_vars

The goal of `dttm_vars` is to compute variables derived from date, such as year, month, day, etc.

```{r dttm_vars}
dt <- seq(as.Date("1910/1/1"), as.Date("1911/1/1"), "days")

df_dt <- data.frame(dt = dt)

dttm_vars(df_dt,dt) %>% dplyr::glimpse()
```

The new variables respective to the date are:

 - **mon_abb:** a factor with the abbreviated name of the month;
 - **mon_lbl:** a factor with the complete name of the month;
 - **mon_num:** the number of the month;
 - **day_num:** the number of the day within the month;
 - **year_num:** the number of the year;
 - **week_num:** the number of complete seven day periods that have occurred between the date and January 1st, plus one;
 - **epiweek_num:** the US CDC version of epidemiological week. Starts on Sunday;
 - **isoweek_num:**  the week as it would appear in the ISO 8601 system, which uses a reoccurring leap week. Starts on Monday;
 - **wday_abb:** a factor with the week day abbreviated name;
 - **wday_lbl:** a factor with the week day complete name.
