---
title: "Functions calc_"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions calc_}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE}
library(relper)
library(dplyr)
library(ggplot2)
```

`calc_` functions compute a certain value.

# calc_acf

The goal of `calc_acf` is to compute the auto-correlation function,
given by:

$$\frac{\sum_\limits{t = k+1}^{n}(x_t - \bar{x})(x_{t-k} - \bar{x})}{\sum_\limits{t = 1}^{n} (x_t - \bar{x})^2 },$$
where:

-   $x_t$ is a time series of length $n$;
-   $x_{t-k}$ is a shifted time series by $k$ units in time;
-   $\bar{x}$ is the average of the time series.

```{r calc_acf}
x <- rnorm(100)

calc_acf(x)

```

If you pass a second vector in the argument `y` the cross-correlation
will be computed instead:

$$\frac{n \left( \sum_\limits{t = 1}^{n}x_ty_t \right) - \left[\left(\sum_\limits{t = 1}^{n}x_t \right) \left(\sum_\limits{t = 1}^{n}y_t\right) \right]}{\sqrt{\left[n \left( \sum_\limits{t = 1}^{n}x_t^2 \right) - \left( \sum_\limits{t = 1}^{n}x_t \right)^2\right]\left[n \left( \sum_\limits{t = 1}^{n}y_t^2 \right) - \left( \sum_\limits{t = 1}^{n}y_t \right)^2\right]}},$$
where:

-   $x_t$ is a time series of length $n$;
-   $y_t$ is a time series of length $n$.

```{r calc_acf-ccf}
y <- rexp(100)

calc_acf(x,y)

```

# calc_association

The goal of `calc_association` is to compute associations metrics.

## Contingency

Contingency is a measure of association between two nominal variables. I
It ranges from 0 and 1, where 0 indicates no relationship and 1
indicates perfect association, and is given by:

$$\sqrt{\frac{X^2}{n+X^2}},$$

where:

-   $X^2$ the chi-square statistic;
-   $n$ is the sample size.

```{r calc_association-contingency}
calc_association(mtcars$am,mtcars$vs,type = "contingency")
```

## Cramér's V

Cramér's V is a measure of association between two nominal variables. I
It ranges from 0 and 1, where 0 indicates no relationship and 1
indicates perfect association, and is given by:

$$\sqrt{\frac{X^2}{n\min(r-1,c-1)}},$$

where:

-   $X^2$ the chi-square statistic;
-   $n$ is the sample size;
-   $r$ is the number of rows in the contingency table;
-   $c$ is the number of columns in the contingency table.

```{r calc_association-cramers-v}
calc_association(mtcars$am,mtcars$vs,type = "cramers-v")
```

## Phi

Phi is a measure of association between two nominal dichotomous
variables, considering a marginal table of the variables given by:

|           | **y = 0** |  **y = 1** | **Total** |
|-----------|-----------|------------|-----------|
| **x = 0** |  $n_{00}$ |  $n_{01}$  |  $n_{0.}$ |
| **x = 1** |  $n_{10}$ |  $n_{11}$  |  $n_{1.}$ |
| **Total** |  $n_{.0}$ |  $n_{.1}$  |  $n$      |

Then the phi coefficient is given by:

$$\frac{n_{11}*n_{00} - n_{10}*n_{01} }{\sqrt{n_{1.}*n_{0.}*n_{.1}*n_{.0}}}.$$
```{r calc_association-phi}
calc_association(mtcars$am,mtcars$vs,type = "phi")
```


# calc_auc

The goal of `calc_auc` is to compute the area under a curve (AUC).

```{r calc_auc-data}
x <- seq(-3,3,l = 100)

y <- dnorm(x)
```

```{r calc_auc-base-plot, echo = F}
ggplot(tibble(x = x,
              y = y),
       aes(x,y))+
  geom_point()+
  plt_theme_y()
```

The function default compute the area considering the range of `x`.

```{r calc_auc-99}
#from min to max of x
range(x)

calc_auc(x,y)
```

```{r calc_auc-plot-99, echo = F}
df_auc <-
  tibble(
    x = x,
    y = y)

df_auc %>% 
ggplot(aes(x,y))+
  geom_area(alpha = .7, fill = "chocolate2")+
  geom_point()+
  plt_theme_y()+
  annotate("text",
           x = mean(x),y = mean(y),label = round(calc_auc(x,y),3),
           fontface = "bold", size = 7)+
  geom_vline(xintercept = c(-3,3), linetype = "dashed")+
  scale_x_continuous(breaks = -5:5)+
  scale_y_continuous(expand = c(0,0))
```

But you can define the argument `limits` to get the AUC of that
respective range.

```{r calc_auc-95}
#from -2 to 2
calc_auc(x,y,limits = c(-2,2))
```

```{r calc_auc-plot-95, echo = F}
df_auc %>% 
  ggplot(aes(x,y))+
  geom_area(data = df_auc %>% 
              filter(between(x,-2,2)),
              alpha = .7, fill = "chocolate2")+
  geom_point()+
  plt_theme_y()+
  annotate("text",
           x = mean(x),y = mean(y),label = round(calc_auc(x,y,limits = c(-2,2)),3),
           fontface = "bold", size = 7)+
  geom_vline(xintercept = c(-2,2), linetype = "dashed")+
  scale_x_continuous(breaks = -5:5)+
  scale_y_continuous(expand = c(0,0))
```

```{r calc_auc-68}
#from -1 to 1
calc_auc(x,y,limits = c(-1,1))
```

```{r calc_auc-plot-68, echo = F}
df_auc %>% 
  ggplot(aes(x,y))+
  geom_area(data = df_auc %>% 
              filter(between(x,-1,1)),
              alpha = .7, fill = "chocolate2")+
  geom_point()+
  plt_theme_y()+
  annotate("text",
           x = mean(x),y = mean(y),label = round(calc_auc(x,y,limits = c(-1,1)),3),
           fontface = "bold", size = 7)+
  geom_vline(xintercept = c(-1,1), linetype = "dashed")+
  scale_x_continuous(breaks = -5:5)+
  scale_y_continuous(expand = c(0,0))
```

# calc_correlation

The goal of `calc_correlation` is to compute associations metrics.

## Kendall

The Kendall correlation coefficient, also called Kendall's Tau coefficient, quantifies the association between two ranked variables.

It was created by Maurice Kendall and is especially beneficial for analyzing non-linear relationships or ranked data. The coefficient is determined by counting the number of concordant pairs (same order of ranks) and discordant pairs (opposite order of ranks) in the data.

$$\frac{n_c-n_d}{\frac{1}{2}*n(n/1)},$$
where:

-   $n_c$ is the number of concordant observations;
-   $n_d$ is the number of discordant observations;
-   $n$ is the number of observations.

```{r calc_correlation-kendall}
calc_correlation(mtcars$hp,mtcars$drat,type = "kendall")
```

## Pearson

The Pearson correlation coefficient is a measure of the linear relationship between two continuous variables. It ranges from -1 to 1, indicating the strength and direction of the association.

A value of 1 represents a perfect positive linear relationship, -1 represents a perfect negative linear relationship, and 0 indicates no linear relationship. 

$$\frac{\sigma_{xy}}{\sigma_x\sigma_y},$$
where:

-   $\sigma_{xy}$ is the covariance of $x$ and $y$;
-   $\sigma_{x}$ is the variance of $x$;
-   $\sigma_{y}$ is the variance of $y$.


```{r calc_correlation-pearson}
calc_correlation(mtcars$hp,mtcars$drat,type = "pearson")
```

## Spearmann

The Spearman correlation coefficient is a measure of the monotonic relationship between two variables, it assesses the strength and direction of the association, regardless of whether it is linear or non-linear. 

It also ranges from -1 to 1, with 1 indicating a perfect monotonic relationship and -1 indicating a perfect inverse monotonic relationship. A value of 0 suggests no monotonic relationship. 

$$1- \frac{6\sum\limits_{i=1}^{n}d_i^2}{n(n^2-1)},$$

where:

-   $d_i$ is the difference between the ranks of $x$ and $y$;
-   $n$ is the number of observations.


```{r calc_correlation-spearman}
calc_correlation(mtcars$hp,mtcars$drat,type = "spearman")
```

# calc_cv

The goal of `calc_cv` is to compute the coefficient of variation (CV),
given by:

$$\frac{s}{\bar{x}},$$ where:

-   $s$ is the sample standard deviation;
-   $\bar{x}$ is the sample mean.

```{r calc_cv}
set.seed(123);x <- rexp(n = 100)

calc_cv(x)
```

If you set the argument `as_perc` to `TRUE`, the CV will be multiplied
by 100.

```{r calc_cv-as-perc}
calc_cv(x,as_perc = TRUE)
```

# calc_geometric_mean

The goal of `calc_geometric_mean` is to compute the geometric mean,
given by:

$$\sqrt[n]{\prod\limits_{i=1}^{n}x_i} = \sqrt[n]{x_1\times x_2 \times...\times x_n},$$
where:

-   $x_i$ is a numeric vector of length $n$.

```{r calc_geometric_mean}
calc_geometric_mean(x)
```

# calc_harmonic_mean

The goal of `calc_harmonic_mean` is to compute the harmonic mean, given
by:

$$\frac{n}{\sum\limits_{i=1}^{n}\frac{1}{x_i}},$$ where:

-   $x_i$ is a numeric vector of length $n$.

```{r calc_harmonic_mean}
calc_harmonic_mean(x)
```

# calc_kurtosis

The goal of `calc_kurtosis` is to compute a kurtosis coefficient.

```{r calc_kurtosis}
calc_kurtosis(x = x)
```

## Biased

The biased kurtosis coefficient, is given by:

$$\frac{\sum\limits_{i=1}^n(x_i - \bar{x})^4}{n*s_x^4},$$

where:

-   $x_i$ is a numeric vector of length $n$;
-   $\bar{x}$ is the mean of $x$;
-   $s_x$ is the standard deviation of $x$.

```{r calc_kurtosis-biased}
calc_kurtosis(x = x,type = "biased")
```

## Excess

The excess kurtosis coefficient, is given by:

$$\frac{\sum\limits_{i=1}^n(x_i - \bar{x})^4}{n*s_x^4}-3,$$

where:

-   $x_i$ is a numeric vector of length $n$;
-   $\bar{x}$ is the mean of $x$;
-   $s_x$ is the standard deviation of $x$.

```{r calc_kurtosis-excess}
calc_kurtosis(x = x,type = "excess")
```
## Percentile

The percentile kurtosis coefficient, is given by:

$$\frac{Q_3-Q_1}{P_{90}-P_{10}},$$ where:

-   $Q_1$ is the first quartile;
-   $Q_3$ is the third quartile;
-   $P_{90}$ is the 90th percentile;
-   $P_{10}$ is the 10th percentile.

```{r calc_kurtosis-percentile}
calc_kurtosis(x = x,type = "percentile")
```

## Unbiased

The unbiased kurtosis coefficient, is given by:

$$\frac{(n+1)*n}{(n-1)*(n-2)*(n-3)}*\frac{\sum\limits_{i=1}^n(x_i - \bar{x})^4}{n*s_x^4} - 3*\frac{(n-1)^2}{(n-2)*(n-3)},$$

where:

-   $x_i$ is a numeric vector of length $n$;
-   $\bar{x}$ is the mean of $x$;
-   $s_x$ is the standard deviation of $x$.

```{r calc_kurtosis-unbiased}
calc_kurtosis(x = x,type = "unbiased")
```

# calc_modality

The goal of `calc_modality` is to compute the number of modes.

```{r calc_modality}

calc_modality(x = c("a","a","b","b"))

```

# calc_mode

The goal of `calc_mode` is to compute the mode.

```{r calc_mode-data}
set.seed(123);cat_var <- sample(letters,100,replace = TRUE)

table(cat_var)

```

We can see that the letter "*y*" is the one that appears the most, so it
is the mode of the variable.

```{r calc_mode}
calc_mode(cat_var)
```

# calc_peak_density

The goal of `calc_peak_density` is to compute the peak density value of
a numeric value.

```{r calc_peak_density-base-plot,echo = F}

pd_plot <-
  ggplot2::ggplot(data = dplyr::tibble(x = x),
                  ggplot2::aes(x))+
  ggplot2::geom_density()+
  relper::plt_theme_y()+
  ggplot2::scale_x_continuous(breaks = 0:20, expand = c(0,0))+
  plt_flip_y_title+
  labs(y = "Density (x)")

pd_plot

```

Let's say we eant to discover what value is the peak in the density
above.

```{r calc_peak_density}
calc_peak_density(x)
```

```{r calc_peak_density-plot,echo = F, warning = FALSE, message = FALSE}
x_peak <- relper::calc_peak_density(x)

pd_plot+
  ggplot2::geom_vline(
    xintercept = relper::calc_peak_density(x),
    col = "royalblue4",
    size = 1
    )+
  ggplot2::scale_x_continuous(
    breaks = 0:20,
    expand = c(0,0),
    sec.axis = sec_axis(~.,
                        breaks = x_peak,
                        labels = format_num(x_peak,digits = 3)
                        )
    )

```

# calc_skewness

The goal of `calc_skewness` is to compute a skewness coefficient.

```{r calc_skewness}
calc_skewness(x = x)
```

Where different types of coefficients are provided, they are:

## Bowley

The Bowley skewness coefficient, is given by:

$$\frac{Q_3+Q_1-2Q_2}{Q_3-Q_1},$$ where:

-   $Q_1$ is the first quartile;
-   $Q_2$ is the second quartile;
-   $Q_3$ is the third quartile.

```{r calc_skewness-bowley}
calc_skewness(x = x,type = "bowley")
```

## Fisher-Pearson

The Fisher-Pearson skewness coefficient, is given by:

$$\frac{\sum_\limits{i=1}^{n}(x_i - \bar{x})^3}{n*(s_x)^3},$$

where:

-   $\bar{x}$ is the mean of $x$;
-   $x_i$ is a numeric vector of length $n$;
-   $s_x$ is the standard deviation of $x$.

```{r calc_skewness-fisher_pearson}
calc_skewness(x = x,type = "fisher_pearson")
```

## Kelly

The Kelly skewness coefficient, is given by:

$$\frac{P_{90}+P_{10}-2Q_2}{P_{90}-P_{10}},$$ where:

-   $P_{90}$ is the 90th percentile;
-   $Q_2$ is the second quartile, i.e., $P_{50}$;
-   $P_{10}$ is the 10th percentile;

```{r calc_skewness-kelly}
calc_skewness(x = x,type = "kelly")
```

## Pearson median

The Pearson median skewness coefficent, or second skewness coefficient,
is given by:

$$\frac{3(\bar{x}- \tilde{x})}{s_x},$$

where:

-   $\bar{x}$ is the mean of $x$;
-   $\tilde{x}$ is the median of $x$;
-   $s_x$ is the standard deviation of $x$.

```{r calc_skewness-pearson_median}
calc_skewness(x = x,type = "pearson_median")
```

## Rao

The Rao skewness coefficient, is given by:

<!-- The Rao's skewness is a measure of skewness that was proposed by C.R. Rao in the paper "Some Measures of Skewness and Kurtosis Based on the Sample Range", published in the journal "The Annals of Mathematical Statistics" in 1948. -->

$$\frac{[n/(n-1)](\bar{x}- \tilde{x})}{\sqrt{(n-2)/n}},$$

where:

-   $\bar{x}$ is the mean of $x$;
-   $\tilde{x}$ is the median of $x$;
-   $n$ is the length of $x$.

```{r calc_skewness-rao}
calc_skewness(x = x,type = "rao")
```
