---
title: "Functions is_"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions is_}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE}
library(relper)
```

`is_` functions returns a boolean, given a condition.

# is_date

The goal of `is_date` is to check if a value is a date.

```{r is_date}
is_date(Sys.time())

is_date(Sys.Date())
```

# is_datetime

The goal of `is_datetime` is to check if a value is a datetime.

```{r is_datetime}
is_date(Sys.time())

is_date(Sys.Date())
```


# is_even

The goal of `is_even` is to check if a value is even.

```{r is_even}
is_even(1)

is_even(2)

is_even(1.1)

is_even(2.2)

```

# is_integer

The goal of `is_integer` is to check if a value is a integer, not considering their variable type but the value.

```{r is_integer}
is_integer(1.1)

is_integer(1)

```

# is_negative

The goal of `is_negative` is to check if a value is negative.

```{r is_negative}
is_negative(1)

is_negative(-1)

```

# is_odd

The goal of `is_odd` is to check if a value is odd.

```{r is_odd}
is_odd(1)

is_odd(2)

is_odd(1.1)

is_odd(2.2)

```

# is_outlier

The goal of `is_outlier` is to check if a value is an outlier, by using the boxplot outlier criteria, given by:

$$ [x < (Q_1 - 1.5 * IQR)] \quad  |  \quad [x > (Q_3 - 1.5 * IQR)],$$

where:

 - $Q_1$ is the first quartile;
 - $Q_3$ is the third quartile;
 - $IQR$ is the interquartile range, e.g., $Q_3-Q_1$.

```{r is_outlier}
x <- c(1,2,3,5,7,8,12,100)

is_outlier(x)

```

# is_positive

The goal of `is_positive` is to check if a value is positive.

```{r is_positive}
is_positive(1)

is_positive(-1)

```

# is_string

The goal of `is_string` is to check if a value is a string, character or factor.

```{r is_string}
is_string("A")

is_string(factor("A"))

```
