---
title: "Functions summary_"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions summary_}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE}
library(relper)
library(dplyr)
```

`summary_` functions summarize data and return metrics related to them.

# summary_cat

The goal of `summary_cat` is to summarize categorical variables.

```{r summary_cat}
set.seed(123);g <- c(sample(letters,100,replace = TRUE),NA)

summary_cat(g)
```

# summary_num

The goal of `summary_num` is to summarize numeric variables.

```{r summary_num}
set.seed(123);x <- c(rnorm(10),NA,10)

summary_num(x) %>% glimpse()
```

It is also possible to return a minimal version of the summary:

```{r summary_num}
summary_num(x,minimal = TRUE) %>% glimpse()
```


# summary_seq

The goal of `summary_seq` is to compute the number of sequential repeated values.

```{r summary_seq}
y <- c(1, 1, 1, 2, 2, 6, 7, 1, 1)

summary_seq(y)

```
