---
title: "Functions as_"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions as_}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE}
library(relper)
library(dplyr)
```

`as_` functions coerce values to another type.

# as_num
The goal of `as_num` is to be a broader version of `as.numeric`, where the input is a string with number and marks, such as "10.000,02".

```{r as_num}
as_num("123.456,78")
```

# as_perc
The goal of `as_perc` is to compute a number as percentage. By default the function will just multiply values by 100.

```{r as_perc}
mtcars %>% 
  count(vs,am) %>% 
  mutate(prop = n/sum(n)) %>% 
  mutate(perc = as_perc(prop))
```

If you set the argument `sum` to `TRUE` the function will divide the values by their total and multiply by 100.

```{r as_perc-sum}
mtcars %>% 
  count(vs,am) %>% 
  mutate(perc = as_perc(n,sum = TRUE))
```
