---
title: "Other functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Other functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE}
library(relper)
library(dplyr)
library(ggplot2)
```

# cut_by_quantile

The goal of `cut_by_quantile` is to divide a numeric variable by a set of quantiles.

```{r cut_by_quantile}

set.seed(123);x <- rnorm(100)

quartiles <- seq(0,1,by = .25)

table(cut_by_quantile(x,q = quartiles))
```

# expand_grid_unique

The goal of `expand_grid_unique` is to create a grid of all combination from two variables, with no repetition of pairs, not matter the position.

```{r expand_grid_unique}
expand_grid_unique(x = 1:3,y = 1:3)
```

You can also set the argument `include_equals` to `TRUE`, so equal pairs are kept.

```{r}
expand_grid_unique(x = 1:3,y = 1:3, include_equals = TRUE)
```

# obj_to_string

The goal of `obj_to_string` is to return the name of an R object as a string.

```{r obj_to_string}
x <- c(1,2,3,5,7,8,12,100)

obj_to_string(x)

```

# row_number_unique

The goal of `row_number_unique` is to get the row number, but considering the unique values of a variable.

```{r row_number_unique}
tibble(x = c(1,1,1,2,3,4,5,5)) %>% 
  mutate(
    row_number = row_number(),
    row_number_unique = row_number_unique(x)
  )
```

# rpearson

The goal of `rpearson` is to simulate data, where two variables will be linear correlated with a normal distribution, using pearson correlation coefficient as an argument.

```{r rpearson}
set.seed(123);df <- rpearson(n = 100, pearson = .85, mean = 3)

df %>% 
  ggplot(aes(x,y))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  plt_theme_xy()
```
