---
title: "str_ functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{str_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(relper)
```

`str_` functions manipulate strings.

# str_clean

The goal of `str_clean` is to remove punctuation and/or accent.

```{r str_clean}
string <- "a..;éâ...íõ"

#remove only punctuation
str_clean(string,remove_accent = FALSE,remove_punct = TRUE)

#remove only accent
str_clean(string,remove_accent = TRUE,remove_punct = FALSE)

#remove both
str_clean(string)
```

# str_keep

The goal of `str_keep` is to keep only a type of character in the string.

```{r str_keep}
string <- "1Aa45Z89$$%#"

str_keep(string,keep = "text")

str_keep(string,keep = "numbers")

str_keep(string,keep = "special")
```

# str_select

The goal of `str_select` is to select part of a string, before, after or between patterns.

```{r str_select}
string <- "example text STRING1 TARGET STRING2 example text"

#Select a string, before a pattern
str_select(string,before = "STRING2")

#Select a string, after a pattern
str_select(string,after = "STRING1")

#Select a string, between two patterns
str_select(string,"STRING1","STRING2")
```

# str_to_abb

The goal of `str_to_abb` is to apply uppercase to strings with a number of characters lower than the argument `n_abb` (default = 3).

```{r str_to_abb}
abcde_string <- c("aaaaa","bb","ccc","dddd")

str_to_abb(abcde_string,n_abb = 3)
```
