---
title: "summary_ functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{summary_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(relper)
```

`summary_` functions summarize data and return metrics related to them.

# summary_cat

The goal of `summary_cat` is to summarize categorical variables.

```{r summary_cat}
x <- c(sample(letters,100,replace = TRUE),NA)

summary_cat(x)
```

# summary_data

The goal of `summary_data` is to summarize all variables from a dataframe.

```{r summary_data}
summary_data(mtcars)
```

# summary_num

The goal of `summary_num` is to summarize numeric variables.

```{r summary_num}
x <- c(rnorm(10),NA,10)

summary_num(x)
```

# summary_seq

The goal of `summary_seq` is to compute the number of sequential repeated values.

```{r summary_seq}
y <- c(1, 1, 1, 2, 2, 2, 2, 3, 4, 5, 6, 7, 1, 1)

summary_seq(y)

```
